FCST

### 2022-05-03 analysis/chapters/forecast_20181122.Rmd
tidying folders.

cordillera <- list(
  ggnewscale::new_scale_fill(),
  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
                                         z = h),
                    breaks = seq(1500, 8000, by = 700)),
  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
                      oob = scales::squish))

@@

experiments <- c("E4", "E5", "E6", "E7")
#Color blind friendly but E4 y E7 parecidos en Desaturated
colores_exp <- c(E4 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"

@@

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[4-7]_ens.csv")

@@

fss %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = factor(ini_date))) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "1 km", "5" = "50 km", 
                                            "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL, linetype = NULL) +
  theme_minimal()

@@

```{r message=FALSE, warning=FALSE}
files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
files <- files[str_detect(files, "_01_")]

sondeos <- purrr::map(files, function(f){

  fread(f) %>% 
    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
    .[, launch_time := as_datetime(launch_time)] %>% 
    .[, fcst_exp := basename(dirname(dirname(f)))]

miembros <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_agregados.rds") %>% 
  .[, mem := formatC(mem, width = 3, flag = "0")] %>% 
  fisica[., on = "mem"] %>% 
  separate(fisica, into = c("cumulus", "pbl"))

media <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_media.rds") %>% 
  .[, .(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
        BIAS = mean(value - fcst_value, na.rm = TRUE)), 
    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est")

@@

sondeos %>% 
  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
           BIAS = mean(value - fcst_value, na.rm = TRUE)), 
    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-media}
extremos <- miembros %>% 
  .[, .(max_rmse = max(RMSE),
        min_rmse = min(RMSE)), by = .(exp, fcst_exp, lev, variable)]

media %>% 
  .[estadistico == "RMSE"] %>% 
  extremos[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_ribbon(aes(ymin = min_rmse, 
                  ymax = max_rmse, fill = exp), alpha = 0.2) +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E7 = "RAD")) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

media %>% 
  .[estadistico == "RMSE"] %>% 
  miembros[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(alpha = 0.1, 
            aes(y = RMSE, color = exp, group = interaction(exp, mem))) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-param}
pbl <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, lev, variable)]

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(pbl = "Mean",
           RMSE = value.est)] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, lev, variable)]

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(cumulus = "Mean",
           RMSE = value.est)] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

@@

```{r bias-sondeos-media}
extremos <- miembros %>% 
  .[, .(max_bias = max(BIAS),
        min_bias = min(BIAS)), by = .(exp, fcst_exp, lev, variable)]

media %>% 
  .[estadistico == "BIAS"] %>% 
  extremos[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_ribbon(aes(ymin = min_bias, 
                  ymax = max_bias, fill = exp), alpha = 0.2) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E7 = "RAD")) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

@@

  geom_line(alpha = 0.1, 
            aes(y = BIAS, color = exp, group = interaction(exp, mem))) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal() geom_hline(yintercept = 0, color = "grey30") +
```

```{r bias-sondeos-param}
pbl <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, lev, variable)]

>>>>

cordillera <- list(
  ggnewscale::new_scale_fill(),
  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
                                         z = h),
                    breaks = seq(1500, 8000, by = 700)),
  scale_fill_gradient(low = "#E2E6E6", high = "#7E8E8E", guide = "none",
                      oob = scales::squish))

@@

experiments <- c("E2", "E5", "E6", "E8")
#Color blind friendly but E4 y E8 parecidos en Desaturated
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE8733", E9 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E8 = "#882255")
satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"

@@

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,8,9]_ens.csv")

@@

fss %>% 
  .[q %in% c(1, 25) & w %in% c(1, 11)] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = factor(ini_date))) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT",
                                                      E6 = "SATWND", E8 = "viejo", E9 = "RAD")) +
  scale_linetype_manual(values = c(1, 2), labels = c("2018-11-22 00:00:00" = "OO UTC", 
                                   "2018-11-22 06:00:00" = "06 UTC")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm", "50" = "50 mm"),
                                      w = c("1" = "10 km", "5" = "50 km", 
                                            "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL, linetype = NULL) +
  theme_minimal()

```{r}
files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2569]_fcst_201811220*pp_pmm.rds")

ppacum_1h <- purrr::map(files, function(f) {

  meta <- unglue(basename(f), "{exp}_fcst_{inidate}_pp_pmm.rds")

  readRDS(f) %>% 
    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
    .[, ":="(rank = NULL, 
             ens_mean = NULL,
             ini_date = ymd_h(meta[[1]][["inidate"]]))] %>% 
  setnames(c("pp_pmm"), c("pp_acum"))

imerg <- readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
  .[x %between% c(-990000, 990000) & y %between% c(-1190000, 1190000)] %>% 
  .[end_date %between% c(as_datetime("20181122060000"), as_datetime("20181123120000"))] %>% 
  .[, exp := "IMERG"]
# %>% 
#   .[, .(pp_acum = mean(pp_acum),
#         lat = mean(lat),
#         exp = "IMERG"), by = .(end_date, y)]

ppacum_1h[, .(mm = sum(pp_acum)/47561), by = .(exp, date, ini_date)] %>% 
  ggplot(aes(date, mm)) +
  geom_line(aes(color =  exp, linetype = factor(ini_date))) +
    scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT",
                                                      E6 = "SATWND", E9 = "RAD")) +
  geom_line(data = imerg[, .(mm = sum(pp_acum)/47561), by = .(exp, end_date)], 
            aes(x = end_date, linetype = "IMERG"), color = "black") +
  labs(x = "Time", y = "Total pp [mm/grid points]", color = NULL, linetype =  NULL) +
  theme_minimal()
```

```{r}
readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
  .[end_date %between% c(as_datetime("20181122060000"), as_datetime("20181123120000"))] %>% 
  .[, .(pp_acum = mean(pp_acum),
        lat = mean(lat),
        exp = "IMERG"), by = .(end_date, y)] %>% 
  ggplot(aes(end_date, lat)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(0.5, 13, 0.5), Inf)) +
  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(0.5, 13, 0.5)) +
  # scale_fill_viridis_c(option = "A",
  scale_fill_distiller(palette = "YlGnBu",
                       direction = 1,
                       labels = function(x) JumpBy(x, 2, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.3, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  scale_date(ini = 20181122060000, break_bin = "12 hours") +
  scale_y_latitude(breaks = seq(-40, -20, 5)) +
  facet_wrap(~ exp) +
  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 9)) +

ppacum_1h %>% 
  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
  .[, lat := lat[1], by = .(y)] %>% 
  .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y, lat)] %>% 
   # .[date %between% c(as_datetime("20181121180000"), as_datetime("20181123120000"))] %>% 
  ggplot(aes(date, lat)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(0.5, 13, 0.5), Inf)) +
  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(0.5, 13, 0.5)) +
  # scale_fill_viridis_c(option = "A",
  scale_fill_distiller(palette = "YlGnBu",
                       direction = 1,
                       labels = function(x) JumpBy(x, 2, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.3, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  scale_date(ini = 20181122060000, break_bin = "12 hours") +
  scale_y_latitude(breaks = seq(-40, -20, 5), labels = NULL) +
  facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
                                                          E6 = "SATWND", E9 = "RAD"))) +
  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 9)) +

  plot_layout(ncol = 2, widths = c(0.2, 0.8), guides = "collect") & theme(legend.position = 'bottom')
```
```{r}

imerg <- read_rds("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds")

q <- c(1, 5, 10, 25, 50)
imerg <- purrr::map(q, function(f){

  imerg %>%
    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
    .[lon %between% lon_band & lat %between% lat_band] %>%
    .[, .(area = sum(pp_acum > f)/.N), by = .(end_date)] %>%
    .[, umbral := f] %>%
    .[]
}) %>%
  rbindlist()

files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2569]_fcst_*_area_acum_1h.rds")

pp_area <- purrr::map(files, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/{exp}_fcst_{ini_date}_area_acum_1h.rds")

  readRDS(f) %>% 
    .[, ":="(ini_date = ymd_h(meta[[1]][["ini_date"]]))]
  # %>% 
  #   .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)]

pp_area %>% 
  .[umbral %in% c(1, 5, 10) ] %>%
  .[, .(mean_area = mean(area),
        max_area = max(area),
        min_area = min(area)), by = .(umbral, date, exp, ini_date)] %>% 
  ggplot(aes(date, mean_area)) +
  geom_line(data = imerg[umbral %in% c(1, 5, 10)], 
            aes(end_date, area, color = "OBS"),
            linetype = 1) +
  geom_line(aes(color = exp, linetype = factor(ini_date))) +
  geom_ribbon(aes(ymin = min_area, ymax = max_area, fill = exp, group = interaction(exp, ini_date)), alpha = .1) +
  scale_color_manual(values = c(OBS = "grey20", colores_exp),
                     labels =  c(E2 = "CONV", E5 = "ASWS", E6 = "SATWND",
                                E9 = "RAD", OBS = "OBS")) +
    scale_fill_manual(values = c(OBS = "grey20", colores_exp),
                 guide = NULL,
                     labels = c(E2 = "CONV", E5 = "ASWS", E6 = "SATWND",
                                E9 = "RAD", OBS = "OBS")) +
  scale_y_continuous(label = scales::percent_format()) +
  scale_x_datetime(limits = c(ymd_h(2018112200), ymd_h(2018112312))) +
  facet_wrap(~umbral, scales = "free_y", 
             labeller = labeller(umbral = c("1" = "1 mm", "5" = "5 mm", 
                                       "10" = "10 mm", "25" = "25 mm"))) +
  labs(y = "Areal coverage", x = "Date", 
       fill = NULL, color = NULL, linetype = "FCST init") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

```{r message=FALSE, warning=FALSE}
# files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E[2,5,6,8]/FCST/*/sondeos/*")
# files <- files[str_detect(files, "_01_")]
# 
# sondeos <- purrr::map(files, function(f){
# 
#   fread(f) %>%
#     .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>%
#     .[, launch_time := as_datetime(launch_time)] %>%
#     .[, fcst_exp := basename(dirname(dirname(f)))]
# 
# }) %>%
#   rbindlist()

miembros <- read_rds("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos_agregados.rds") %>% 
  .[, mem := formatC(mem, width = 3, flag = "0")] %>% 
  fisica[., on = "mem"] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  setDT() %>% 
  .[, BIAS := -BIAS]

media <- read_rds("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos_media.rds") %>% 
  .[, .(RMSE = sqrt(mean((fcst_value - value)^2, na.rm = TRUE)), 
        BIAS = mean(fcst_value - value, na.rm = TRUE)), 
    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est")

@@

media %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip(xlim = c(0, 15)) +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-ensamble}
extremos <- miembros %>% 
  .[, .(max_rmse = max(RMSE),
        min_rmse = min(RMSE)), by = .(exp, fcst_exp, lev, variable)]

media %>% 
  .[estadistico == "RMSE"] %>% 
  extremos[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_ribbon(aes(ymin = min_rmse, 
                  ymax = max_rmse, fill = exp), alpha = 0.2) +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E8 = "RAD")) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip(xlim = c(0,  15)) +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

media %>% 
  .[estadistico == "RMSE"] %>% 
  miembros[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(alpha = 0.1, 
            aes(y = RMSE, color = exp, group = interaction(exp, mem))) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-param}
pbl <- miembros[, .(RMSE = mean(RMSE),
                    tipo = "pbl"), 
                by = .(exp, fcst_exp, pbl, lev, variable, iop)] %>% 
  setnames("pbl", "param")

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("iop", "exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           RMSE = value.est, 
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip() +
  facet_grid(factor(fcst_exp) +iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(RMSE = mean(RMSE),
                    tipo = "cumulus"), 
                by = .(exp, fcst_exp, cumulus, lev, variable, iop)] %>% 
  setnames("cumulus", "param")

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("iop", "exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           RMSE = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip() +
  facet_grid(factor(fcst_exp) + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

rbind(pbl, cum) %>% 
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp)] %>% 
  ggplot(aes(RMSE_mean)) +
  geom_histogram(aes(fill = exp), 
                 position = "identity", 
                 alpha = 0.5) +
   scale_fill_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "AUT",
                                E6 = "SATWND", E8 = "RAD")) +
  facet_grid(fcst_exp ~ variable, scales = "free_x")

rbind(pbl, cum) %>%
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(RMSE_mean, exp)) +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E2 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E8 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E2 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E8 = "RAD")) +
  facet_grid(fcst_exp ~variable, scales = "free_x") +
  labs(y = NULL, x = "mean RMSE", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))

@@

```{r bias-sondeos-ensamble}
extremos <- miembros %>% 
  .[, .(max_bias = max(BIAS),
        min_bias = min(BIAS)), by = .(exp, fcst_exp, lev, variable)]

media %>% 
  .[estadistico == "BIAS"] %>% 
  extremos[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_ribbon(aes(ymin = min_bias, 
                  ymax = max_bias, fill = exp), alpha = 0.2) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E8 = "RAD")) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

@@

  geom_line(alpha = 0.1, 
            aes(y = BIAS, color = exp, group = interaction(exp, mem))) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal() 
```

```{r bias-sondeos-param}
pbl <- miembros[, .(BIAS = mean(BIAS),
                    tipo = "pbl"), 
                by = .(exp, fcst_exp, pbl, lev, variable)] %>% 
  setnames("pbl", "param")

media %>% 
  .[estadistico == "BIAS"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           BIAS = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, BIAS)) +
  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(BIAS = mean(BIAS),
                    tipo = "cumulus"), 
                by = .(exp, fcst_exp, cumulus, lev, variable)] %>% 
  setnames("cumulus", "param")
